---
title: "02_tree_structure"
author: "randy"
date: '2022-04-03'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction to phylogenies in R

The codes are from 

[Martin R. Smith](https://cran.r-project.org/web/packages/TreeTools/vignettes/load-trees.html)
and 
[Liam J. Revell.](http://www.phytools.org/eqg/Exercise_3.2/)
and 
[Rileigh Dowling](https://rstudio-pubs-static.s3.amazonaws.com/675027_20485b2a006a49ba9a07edb5034c37cd.html)


## phylo structure

```{r eval=FALSE, include=FALSE}
## install CRAN Task View for phylogenetics
#  CRAN are relevant for tasks related to a certain topic.
install.packages('ctv')
library(ctv) 
update.views('Phylogenetics')
install.views('Phylogenetics')
## update.views('Phylogenetics')
```
```{r fig.height=5, fig.width=5}
library(ape)
## simulate a phylogeny
tree <- rtree(n = 28)
plot(tree, edge.width = 2)

# str(tree)
# View(tree)
```


a list of class "phylo"

- the core of how an object of class "phylo" encodes phylogenetic

- let's use a simple case: a tree with 5 tips & no edge lengths


```{r fig.height=2, fig.width=2}
tree <- read.tree(text = "(((A, B), (C, D)), E);")
plot(tree, 
     type = "cladogram", 
     edge.width = 2)
```

```{r}
tree$edge
tree$tip.label
tree$Nnode
```

```{r fig.height=2, fig.width=2}
plot(tree, 
     edge.width = 2, 
     label.offset = 0.1, 
     type = "cladogram")
nodelabels()
tiplabels()
```

- all of the relationship information among the taxa in the tree is containing in the starting & ending nodes for each edge

- edges that share a common starting node number are descended from an immediate common ancestor

- object of class "phylo"

- useMethods with plot.phylo

- storing trees in the same way across different packages for  interoperability


## Writing & reading phylogenetic trees

```{r}
## write a tree file and save it
write.tree(tree, "data/example.tre")
cat(readLines("data/example.tre"))

## load phytools
library(phytools)
writeNexus(tree, "data/example.nex")
cat(readLines("data/example.nex"), sep = "\n")
```

## Simulating, plotting, extracting clades, & dropping tips


```{r fig.height=4, fig.width=4}
set.seed(1)
## simulate a birth-death tree using phytools
?pbtree 
               ##  birth rate or speciation rate 
tree <- pbtree(b = 1, 
               ## death rate or extincition rate 
               d = 0.2, 
               ## sample size for simulation
               n = 40)

plotTree(tree, setEnv = TRUE)
nodelabels()
tiplabels()
```

```{r fig.height=3, fig.width=3}
## extract the descendeed tree from #62
tt62 <- extract.clade(tree, 62)
plotTree(tt62)
nodelabels()
tiplabels()
```


```{r fig.height=3, fig.width=3}
## now drop 10 tips from the tree 
## (I'm going to pick them at random)
dtips <- sample(tree$tip.label, 10)
dt <- drop.tip(tree, dtips)

## now plot the cut tree
plotTree(dt)
nodelabels()
tiplabels()
```


```{r fig.height=8, fig.width=4}
## we could also, say, 
## drop all tips that go extinct before the present
## this is a fun way, 
## but not the only way to do this:
et <- fancyTree(tree, 
                type = "droptip", 
                tip = getExtinct(tree),
                cex = 0.7)
nodelabels()
tiplabels()
```

```{r}
## now this is the extinction tree
print(et)
View(et)
```


# Binary & polytomous trees

"ape" and most other phylogenetics packages are equipped to handle phylogenies that are binary or multifurcating.

```{r fig.height=2, fig.width=2}
t1 <- read.tree(text = "((A, B, C), D);")
plot(t1, type = "cladogram")

## check whether it is a binary tree 
is.binary.tree(t1)  

## randomly resolve polytomies
t2 <- multi2di(t1)
plot(t2, type = "cladogram")
```

## Miscellaneous 


```{r fig.height=4, fig.width=4}
## this is the original tree
plotTree(tree, node.numbers = T)

op <- par(mfrow = c(2, 2))

## first, rotate about node #50
rt.50 <- rotate(tree, 50)
plotTree(rt.50)

## now rotate all nodes
rt.all <- rotateNodes(tree, "all")
plotTree(rt.all)

## let's re-root the tree at node #67
rr.67 <- root(tree, node = 67)
plotTree(rr.67)

## this creates a trifurcation 
## at the root we could instead re-root at
## along an edge
rr.62 <- reroot(tree, 62, position = 0.5 * tree$edge.length[which(tree$edge[, 
    2] == 62)])
plotTree(rr.62)

par(op)
```

Comparing trees

```{r}
## check if tree & rt.all are equal
all.equal(tree, rt.all)
## [1] TRUE


## check if tree & rr.62 are equal
all.equal(tree, rr.62)
## [1] FALSE


## check if unrooted tree & rr.62 are equal
all.equal(unroot(tree), unroot(rr.62))
```

Generating multiple trees

```{r fig.height=4, fig.width=4}
library(tidyverse)

## simulate 10 pure-birth trees
trees <- pbtree(n = 6, nsim = 9, scale = 1)
op <- par(mfrow = c(3, 3))
## 10 phylogenetic trees
trees %>% map(plot)
par(op)


## round the edge lengths of the tree to 3 digits
trees <- roundBranches(trees, 1)
## write to file
write.tree(trees, file = "data/example.trees")
cat(readLines("data/example.trees"), sep = "\n")
```

# Building a distance-based tree 

```{r fig.height=2, fig.width=4}
library(phangorn)

set.seed(55)
p <- 10

A <- matrix(runif(p ^ 2) * 2 - 1, 
            ncol = p) 
M <- t(A) %*% A %>%
  cov2cor()

# ## now build a matrix
# M <- matrix(runif(p * p, 0, 1),
#             nrow = p,
#             byrow = T)
# diag(M) <- 1
# # View(lower.tri(M))
# 
# ## make the lower triangle 
# ## the same as the upper
# M[lower.tri(M)] <- t(M[upper.tri(M)])
# View(M)

row.names(M) <- LETTERS[1:p]
colnames(M) <- LETTERS[1:p]

## this is the dis-similarity matrix
D <- 1 - M

## convert to distrance format
Dist <- as.dist(D)
str(Dist)

## build a neighbor - joining tree
Tnj <- ape::nj(Dist)
# ?nj # Neighbor-Joining Tree Estimatio

op <- par(mfrow = c(1, 2))

plot(Tnj, "unrooted")
## plot with a root
plot(Tnj)

par(op)
```

Neighbor Joining is one of the most common ways 
to build a tree using molecular data 
that’s been converted to sequences.

its one of the options within BLAST.

```{r fig.height=2, fig.width=2}
Upgma <- phangorn::upgma(Dist)
# View(Upgma)
plot(Upgma)
```

