---
title: "01_creat_package"
author: "randy"
date: '2022-04-24'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
# install.packages(c("devtools", 
#                    "roxygen2", 
#                    "testthat", 
#                    "usethis", 
#                    "knitr"))

library(testthat)
library(usethis)
library(devtools)
```



```{r}
packageVersion("devtools")
# create_package("~/Desktop/bios7732/bios7732_package/LDTM")
```

```{r}

R.home()
list.files()
# list.dirs()
list.dirs(path = ".",
          full.names = FALSE,
          recursive = FALSE)
```

RStudio has special handling for packages 
and you should now see a **Build** tab 
in the same pane as Environment and History.


In the file browser, go to `More > Show Hidden Files` to toggle the visibility of hidden files (a.k.a. “dotfiles”). A select few are visible all the time, but sometimes you want to see them all.

- .Rbuildignore lists files that we need to have around but that should not be included when building the R package from source.


- .Rproj.user, if you have it, is a directory used internally by RStudio.

- .gitignore anticipates Git usage and ignores some standard, behind-the-scenes files created by R and RStudio. Even if you do not plan to use Git, this is harmless.

- DESCRIPTION provides metadata about your package. We edit this shortly.

- NAMESPACE declares the functions your package exports for external use and the external functions your package imports from other packages. At this point, it is empty, except for a comment declaring that this is a file we will not edit by hand.

- The R/ directory is the “business end” of your package. It will soon contain .R files with function definitions.

- package.Rproj is the file that makes this directory an RStudio Project. Even if you don’t use RStudio, this file is harmless. Or you can suppress its creation with create_package(..., rstudio = FALSE). 

```{r}
use_git()
```

