---
title: "01_creat_package"
author: "randy"
date: '2022-04-24'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
# install.packages(c("devtools", 
#                    "roxygen2", 
#                    "testthat", 
#                    "usethis", 
#                    "knitr"))
library(tidyverse)
library(testthat)
library(usethis)
library(devtools)
```

```{r}
packageVersion("devtools")
## to create a package
# create_package("~/Desktop/bios7732/bios7732_package/LDTM")
```

```{r}
## look for R home directory
R.home()
## look at the files in the same directory
here::here() %>%
  list.files()
# list.dirs()
here::here() %>%
  list.dirs(full.names = FALSE,
            recursive = TRUE)
```

RStudio has special handling for packages 
and you should now see a **Build** tab 
in the same pane as Environment and History.


In the file browser, go to `More > Show Hidden Files` to toggle the visibility of hidden files (a.k.a. “dotfiles”). A select few are visible all the time, but sometimes you want to see them all.

- .Rbuildignore lists files that we need to have around but that should not be included when building the R package from source.


- .Rproj.user, if you have it, is a directory used internally by RStudio.

- .gitignore anticipates Git usage and ignores some standard, behind-the-scenes files created by R and RStudio. Even if you do not plan to use Git, this is harmless.

- DESCRIPTION provides metadata about your package. We edit this shortly.

- NAMESPACE declares the functions your package exports for external use and the external functions your package imports from other packages. At this point, it is empty, except for a comment declaring that this is a file we will not edit by hand.

- The R/ directory is the “business end” of your package. It will soon contain .R files with function definitions.

- package.Rproj is the file that makes this directory an RStudio Project. Even if you don’t use RStudio, this file is harmless. Or you can suppress its creation with create_package(..., rstudio = FALSE). 


```{r}
use_git()
```


```{r}
## example for a function
(x <- "alfa,bravo,charlie,delta")
strsplit(x, split = ",")

## This makes total sense in light of R’s 
## fundamental tendency towards vectorization. 
## But sometimes it’s still a bit of a bummer
str(strsplit(x, split = ","))

## The second, safer solution is the basis 
## for the inaugural function
strsplit1 <- function(x, split) {
  strsplit(x, split = split)[[1]]
}
```


Packages and scripts use different mechanisms to declare their dependency on other packages and to store example or test code.

```{r}
use_r("01_tree")
use_r("02_likelihood")
use_r("03_dirichlet")
use_test("01_tree")
use_test("02_likelihood")
use_test("03_dirichlet")
```


If this were a regular R script, 
we might use RStudio to send the function definition to
the R Console and define strsplit1() in the global environment. 
Or maybe we’d call source("R/strsplit1.R").

```{r}
load_all()
exists("Ytree",
       # where = globalenv(), 
       inherits = FALSE)
#> [1] FALSE

# ?exists Environment Access
env <- environment(Ytree)
exists("Btree", 
       where = env,
       # where = globalenv(), 
       inherits = FALSE)

```


This may seem silly to check, after such a small addition, but it’s good to establish the habit of checking this often.

```{r}
check()
```













