
```{r}
usethis::use_tidy_description()
load_all()
```


```{r}
usethis::use_package("dplyr") # Default is "Imports"

usethis::use_package("ggplot2", "Suggests")
requireNamespace("ggplot2", quietly = TRUE)
```


```{r}
# the suggested package is required 
my_fun <- function(a, b) {
  if (!requireNamespace("pkg", quietly = TRUE)) {
    stop(
      "Package \"pkg\" must be installed to use this function.",
      call. = FALSE
    )
  }
  # code that includes calls such as pkg::f()
}

# the suggested package is optional; a fallback method is available
my_fun <- function(a, b) {
  if (requireNamespace("pkg", quietly = TRUE)) {
    pkg::f()
  } else {
    g()
  }
}
```


```{r}
# the suggested package is required 
my_fun <- function(a, b) {
  rlang::check_installed("pkg", reason = "to use `my_fun()`")
  # code that includes calls such as pkg::f()
}

# the suggested package is optional; a fallback method is available
my_fun <- function(a, b) {
  if (rlang::is_installed("pkg")) {
    pkg::f()
  } else {
    g()
  }
}

```


These rlang functions have handy features for programming, 
such as vectorization over pkg, 
classed errors with a data payload, and, for check_installed(), 
an offer to install the needed package in an interactive session.


```{r}
#' @examples
#' if (require("maps")) {
#'   nz <- map_data("nz")
#'   # Prepare a map of NZ
#'   nzmap <- ggplot(nz, aes(x = long, y = lat, group = group)) +
#'     geom_polygon(fill = "white", colour = "black")
#'
#'   # Plot it in cartesian coordinates
#'   nzmap
#' }
```


An example is basically the only place where we would use `require()` inside a package.

Another place you might use a suggested package is in a vignette.

The tidyverse team generally writes vignettes as if all suggested packages are available.
But if you choose to use suggested packages conditionally in your vignettes, 
the knitr chunk options purl and eval may be useful for achieving this. 


```{r}
check()
use_test("04_simulation")
```


```{r}
# rm(list = c("simulate_DM", "simulate_DMLM", "simulate_DTM"))
devtools::load_all()
check()
```


